Non-kissing and non-crossing complexes \\ for gentle algebras

### 2018-07-02 draft/accordionComplexSurfaces.tex
ajout d'un cas dans la demo accoreons = slalom\dual
general walks

We consider a \defn{bound quiver}~$\bar Q = (Q,I)$, formed by a quiver~$Q$ and an admissible ideal~$I$ of the path algebra~$kQ$, so that the quotient algebra~$kQ/I$ is finite dimensional.

@@

\begin{definition}
\label{def:walk}
A \defn{walk} of a gentle quiver~$\bar Q$ is a maximal string of its blossoming quiver~$\bar Q \blossom$ (thus joining two blossom vertices of~$\bar Q \blossom$).
We denote by~$\walks(\bar Q)$ the set of all walks on~$\bar Q$.
As for strings, we implicitly identify the two inverse walks~$\omega$ and~$\omega^{-1}$, and we let~$\walks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \walks(\bar Q)}$ denote the set of \defn{undirected walks} of~$\bar Q$.
\end{definition}

@@

\begin{definition}
\label{def:accordion}
A \defn{$\dissection$-accordion} is a $B$-curve~$\alpha$ of~$\bar\surface$ such that whenever~$\alpha$ meets a face~$f$~of~$\dissection$,
\begin{itemize}
\item it enters crossing an edge~$e_1$ and leaves crossing an edge~$e_2$ (in other words, $\alpha$ is not allowed to circle around~$f\dual$ when~$f\dual$ is a puncture),
\item the two edges~$e_1$ and~$e_2$ of~$f$ crossed by~$\alpha$ are adjacent on the boundary of~$f$,
\item $\alpha$, $e_1$ and~$e_2$ bound a disk inside~$f$ that does not contain~$f\dual{}$.
\end{itemize}
\end{definition}

@@

In this section, we show that non-kissing complexes and non-crossing complexes actually coincide.
Namely, any pair of dual cellular dissections~$(\dissection, \dissection\dual)$ defines a gentle bound quiver~$\bar Q_{\dissection, \dissection\dual}$ such that the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
Conversely, any gentle bound quiver~$\bar Q$ gives rise to a surface~$\surface_{\bar Q}$ equipped with a pair of dual cellular dissections~$\dissection_{\bar Q}$ and~$\dissection\dual_{\bar Q}$ such that the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, \dissection\dual_{\bar Q}]$.
In fact, we show that the accordion, slalom, and non-kissing complexes can all be pictured in a unique way on the surface.
Interestingly, via the correspondence between pairs of dual cellular dissections and gentle bound quivers, the duality between the dissections~$\dissection$ and~$\dissection\dual$ translates to the Koszul duality between the gentle algebras of~$\bar Q$ and~$\bar Q\dual$.

@@

\begin{definition}
\label{def:surfaceQuiver}
The surface~$\surface(\bar Q)$ of a gentle bound quiver~$\bar Q = (Q,I)$ is obtained by glueing lozanges around arrows of~$Q$ following the rule given by the relations~$I$. The construction also automatically defines two dual cellular dissections~$\dissection_{\bar Q}$ and~$\dissection\dual_{\bar Q}$ of the surface~$\surface_{\bar Q}$.
\end{definition}

\begin{proposition}
\label{prop:dualityKoszul2}
For any gentle bound quiver~$\bar Q$ with Koszul dual~$\bar Q\dual$, the surfaces~$\surface_{\bar Q}$ and~$\surface_{\bar Q\dual}$ coincide, but~$\dissection_{\bar Q\dual} = \dissection\dual_{\bar Q}$ and~$\dissection\dual_{\bar Q\dual} = \dissection_{\bar Q}$.
\end{proposition}

\subsection{Accordion, slalom and non-kissing complexe coincide}

@@

\begin{lemma}
\label{lem:nonKissing=nonCrossing}
Non-kissing walks = non-crossing arcs.
\end{lemma}

\begin{theorem}
\label{thm:complexesCoincide}
\begin{itemize}
\item For any gentle bound quiver~$\bar Q$, the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, \dissection\dual_{\bar Q}]$, therefore to the accordion complex~$\AC[\dissection_{\bar Q}]$ and to the slalom complex~$\SC[\dissection\dual_{\bar Q}]$.
\item For any pair of dual cellular dissections~$\dissection, \dissection\dual$ of an oriented surface, the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
\end{itemize}
\end{theorem}

>>>>

%We consider a \defn{bound quiver}~$\bar Q = (Q,I)$, formed by a quiver~$Q$ and an admissible ideal~$I$ of the path algebra~$kQ$, so that the quotient algebra~$kQ/I$ is finite dimensional.
We consider a \defn{bound quiver}~$\bar Q = (Q,I)$, formed by a quiver~$Q$ and an admissible ideal~$I$ of the path algebra~$kQ$.
Note that we \textbf{do not require} that the quotient algebra~$kQ/I$ be finite dimensional.

@@

\begin{definition}
\label{def:walk}
A \defn{walk} of a gentle quiver~$\bar Q$ is a maximal string of its blossoming quiver~$\bar Q \blossom$, meaning that at each end it either reaches a blossom vertex of~$\bar Q \blossom$ or enters an infinite cycle.
We denote by~$\walks(\bar Q)$ the set of all walks on~$\bar Q$.
As for strings, we implicitly identify the two inverse walks~$\omega$ and~$\omega^{-1}$, and we let~$\walks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \walks(\bar Q)}$ denote the set of \defn{undirected walks} of~$\bar Q$.
\end{definition}

@@

\begin{definition}
\label{def:accordion}
A \defn{$\dissection$-accordion} is a $B$-curve~$\alpha$ of~$\bar\surface$ such that whenever~$\alpha$ meets a face~$f$~of~$\dissection$,
\begin{itemize}
\item it enters crossing an edge~$e_1$ and leaves crossing an edge~$e_2$ (in other words, $\alpha$ is not allowed to circle around~$f\dual$ when~$f\dual$ is a puncture),
\item the two edges~$e_1$ and~$e_2$ of~$f$ crossed by~$\alpha$ are consecutive on the boundary of~$f$,
\item $\alpha$, $e_1$ and~$e_2$ bound a disk inside~$f$ that does not contain~$f\dual{}$.
\end{itemize}
\end{definition}

@@

In this section, we show that non-kissing complexes and non-crossing complexes actually coincide.
Namely, any pair of dual cellular dissections~$(\dissection, \dissection\dual)$ defines a gentle bound quiver~$\bar Q_{\dissection, \dissection\dual}$ such that the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
Conversely, any gentle bound quiver~$\bar Q$ gives rise to a surface~$\surface_{\bar Q}$ equipped with a pair of dual cellular dissections~$\dissection_{\bar Q}$ and~${\dissection_{\bar Q}}\dual$ such that the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$.
In fact, we show that the accordion, slalom, and non-kissing complexes can all be pictured in a unique way on the surface.
Interestingly, via the correspondence between pairs of dual cellular dissections and gentle bound quivers, the duality between the dissections~$\dissection$ and~$\dissection\dual$ translates to the Koszul duality between the gentle algebras of~$\bar Q$ and~$\bar Q\dual$.

@@

\begin{definition}
\label{def:surfaceQuiver}
The surface~$\surface(\bar Q)$ of a gentle bound quiver~$\bar Q = (Q,I)$ is obtained by glueing lozanges around arrows of~$Q$ following the rule given by the relations~$I$. The construction also automatically defines two dual cellular dissections~$\dissection_{\bar Q}$ and~${\dissection_{\bar Q}}\dual$ of the surface~$\surface_{\bar Q}$.
\end{definition}

\begin{proposition}
\label{prop:dualityKoszul2}
For any gentle bound quiver~$\bar Q$ with Koszul dual~$\bar Q\koszul$, the surfaces~$\surface_{\bar Q}$ and~$\surface_{\bar Q\koszul}$ coincide, but~$\dissection_{\bar Q\koszul} = {\dissection_{\bar Q}}\dual$ and~${\dissection_{\bar Q\koszul}}\dual = \dissection_{\bar Q}$.
\end{proposition}

\subsection{Non-crossing and non-kissing complexe coincide}

@@

\begin{lemma}
\label{lem:nonKissing=nonCrossing}
Non-kissing walks = non-crossing $\dissection$-accordion curves = non-crossing $\dissection\dual$-slalom curves.
\end{lemma}

\begin{theorem}
\label{thm:complexesCoincide}
\begin{itemize}
\item For any gentle bound quiver~$\bar Q$, the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$, therefore to the accordion complex~$\AC[\dissection_{\bar Q}]$ and to the slalom complex~$\SC[{\dissection_{\bar Q}}\dual]$.
\item For any pair of dual cellular dissections~$\dissection, \dissection\dual$ of an oriented surface, the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
\end{itemize}
\end{theorem}

\section{Pseudomanifolds}

In this section, we show that the non-crossing and non-kissing complexes are \defn{pseudomanifold}, \ie that there are
\begin{enumerate}[(i)]
\item \defn{pure}: all maximal $\dissection$-accordion simplices have the same number of $B$-curves as~$\dissection$, and
\item \defn{thin}: any codimension~$1$ simplex of~$\AC$ is contained in exactly two maximal $\dissection$-accordion simplices.
\end{enumerate}

\vincent{Peut-on utiliser une projection d'une alg\`ebre ammass\'ee ? Apparement non... On peut avoir le Koszul dual d'une alg\`ebre ammass\'ee.}
